---
title: "Jags Robust Design Verification"
author: "M.E. Colvin"
date: "2017-07-26"
output:
  html_document
bibliography: ref.bib
csl: ecology.csl  
---

This code simulates a population and capture histories that can be 
estimated using a robust design. The code is overkill because it allows
temporary emigration among secondary occasions, a big no, no but if 
you set the daily migration rate to 0 it is the base robust design.
The function takes a list of inputs:

* Number of primary occasions
* Survival, Phi
* Capture probability
* Movement among primary occasions
* Movement among secondary occasions
* Recruitment
   
```{r}
sim_ch<-function(inputs,...)
    {library(rowr)
    ## SET UP TIME INTERRVALS
    ## FOR PROGRAM MARK
    ends<-cumsum(inputs$nsec) # last sampling occasion
    occs<- rep(0,sum(inputs$nsec))
    occs[ends]<-1# last occasion in primary
    occs<- occs[-length(occs)]# drop last 1 for processing

    # SET UP SURVIVAL
    S<-occs
    S[which(S==1)]<-inputs$phi
    S[which(S==0)]<-1
    
    # SET UP MOVEMENT
    ## USES GAMMA DOUBLE PRIME PROBABILITY OF BEING IN OUTSIDE STUDY AREA
    ## probability of temporarily emigrating from the observable sample during
    ## an interval is the same as the probability of staying away
    GammaDblPrime<-occs    
    GammaDblPrime[which(GammaDblPrime==1)]<-inputs$gam_d_prime
    GammaDblPrime[which(GammaDblPrime==0)]<-unlist(lapply(1:inputs$nprim,
        function(x) {rep(inputs$gam_d_prime2[x],inputs$nsec[x]-1)}))

    # SET UP RECRUITMENT
    f<- occs
    f[which(f==1)]<-inputs$f
    f[which(f==0)]<-0
  
    ## MAKE THE SUPER POPULATION AT T=1
    Z<- matrix(0,inputs$n,sum(inputs$nsec))
    Z_inn<-Z
    Z_out<-Z
    Z[,1]<-1   
    Z_inn[,1]<-rbinom(nrow(Z),1, 1-inputs$gam_d_prime[1])   
    Z_out[,1]<-1-Z_inn[,1]  
    
    ## SIMULATE POPULATION DYNAMICS
    ### SIMULATES PRIMARY AND SECONDARY OCCASIONS...
    for(i in 2:ncol(Z))
        {
        ## SURVIVAL
        Z[,i]<- rbinom(n=nrow(Z),
            size=1,
            prob=S[i-1]*Z[,i-1])
        ## MOVEMENT: RANDOM
        if(GammaDblPrime[i-1]==0)
            {
            Z_inn[,i]<- Z_inn[,i-1]*Z[,i]
            }
        if(GammaDblPrime[i-1]>0)
            {
            Z_inn[,i]<- rbinom(nrow(Z),1,1-GammaDblPrime[i-1])*Z[,i]
            }
        Z_out[,i]<- (1-Z_inn[,i])*Z[,i]      
        
        ## RECRUITMENT
        ## ASSUME RECRUITS SETTLE OUTSIDE OF 
        ## STUDY AREA
        R<-rpois(1,sum(Z[,i])*f[i-1])
        app<- matrix(0,R,ncol(Z))
        app[,i]<-1
        Z<- rbind(Z,app)
        Z_out<- rbind(Z_out,app)
        app[]<-0
        Z_inn<- rbind(Z_inn,app)
        }

    ## SIMULATE CAPTURE HISTORIES
    ch<- matrix(0,nrow(Z),ncol(Z))
    p<- rep(inputs$p,inputs$nsec)
    for(i in 1:sum(inputs$nsec))
        {
        ch[,i]<-rbinom(nrow(Z),1,p[i]*Z_inn[,i])
        }
       
    ## SUBSET OUT FISH THAT ARE NEVER CAPTURED
    ch<- ch[which(apply(ch,1,sum)!=0),]
    ## NEED THIS FOR RMARK
    # prep data for processing
    ch1<- data.frame(ch=apply(ch,1,paste0,collapse=""),
        freq=1,stringsAsFactors=FALSE)
    ch_array<- 
    out<-(list(ch=ch1,
        ch_mat=ch,
        occs=occs,
        nprim=inputs$nprim,
        nsec=inputs$nsec,
        Z=Z,
        trueN=colSums(Z_inn[,c(which(occs==1),length(occs))]),
        Z_inn=Z_inn,
        Z_out=Z_out))
    return(out)
    }
```

## The BUGS model

The function below contains the BUGS model used to 
estimate the demographic parameters and abundance. It was 
modified from [@RN5251]

```{r}
mod_RD<- function()
	{
        
	###################################################################	
	# PRIORS
    ###################################################################
    lo_omega~dnorm(0,0.37)# INCLUSION PROBABILITY (LOG ODDS)
    lo_gamma~dnorm(0,0.37)# PROBABLITY OF MOVING  (LOG ODDS)
    lo_phi~dnorm(0,0.37) # log odds phi
    lo_p~dnorm(0,0.37) # log odds p
        
    for(d in 1:10)
        {        
        logit(omega[d])<- lo_omega
        logit(S[d])<- lo_phi
        logit(gammaPrime[d])<- lo_gamma
        logit(gammaPrimePrime[d])<- lo_gamma
        logit(p[d])<-lo_p
        lambda[d] <- (1-gammaPrime[d])/(gammaPrimePrime[d]-gammaPrime[d]+1) 
        }

    for(k in 1:10)
        {
        #alive_i[i,t_] <- step(z[i,t_]-3) # check alive or not
        #Nin_i[i,t_] <- equals(z[i,t_],4) # count if i is within study area
        
        # recruitment process from 'eigenvector decomposition'
        # recruitment ratio, or long-term prob of being inside     

        
        # TRANSITION MATRIX
        # 1 =not yet in population;2=dead;3=offsite;4=onsite (only observable state)
        # transition are from the column --> rows

        tmat[1,1,k]<- 1-omega[k] # not in population yet
        tmat[2,1,k]<- 0
        tmat[3,1,k]<- omega[k]*lambda[k]
        tmat[4,1,k]<- omega[k]*(1-lambda[k])
        
        # DEAD
        tmat[1,2,k]<-0
        tmat[2,2,k]<-1 # STAY DEAD
        tmat[3,2,k]<-0
        tmat[4,2,k]<-0
        # OFFSITE
        tmat[1,3,k]<-0
        tmat[2,3,k]<-1-S[k] # DIES 
        tmat[3,3,k]<-0.7#gammaPrime[k]*omega[k] # outside -> outside
        tmat[4,3,k]<-0.7#(1-gammaPrime[k])*omega[k] # outside -> inside
        # ONSITE
        tmat[1,4,k]<- 0
        tmat[2,4,k]<- 1-S[k]
        tmat[3,4,k]<- gammaPrimePrime[k]*omega[k] # inside -> outside
        tmat[4,4,k]<- (1-gammaPrimePrime[k])*omega[k] # inside-> inside
        }


    ###################################################################
    # THE MODEL    
    ###################################################################    
    for(m in 1:M)
        {
        z[m,1]~ dcat(tmat[1:4,1,1]) 
        for(j in 1:T2[1])
            {
            y[m,j,1] ~ dbern(p[1]*equals(z[m,1],4)) 
            }    
        for(i in 2:T)
            {
            z[m,i] ~ dcat(tmat[1:4, z[m,i-1],i])
            for(j in 1:T2[i])
                {
                y[m,j,i] ~ dbern(p[i]*equals(z[m,i],4)) 
                } 
            }
        }        
    }
```

```{r}
input<-list()
input$nprim = 10
input$phi=rep(0.8,input$nprim-1)## SURVIVAL
# FECUNDITY
input$f=rep(0,input$nprim-1)
input$n=500 # initial population size
## RANDOM MOVEMENT PRIMARY
input$gam_d_prime=rep(0.3,input$nprim-1) # OBSERVABLE[t-1] --> UNOBSERVABLE[t]; TEMP EMIGRATION
input$nsec=rep(4,input$nprim)## SECONDARY OCCASIONS
## RANDOM MOVEMENT SECONDARY
input$gam_d_prime2<-rep(0,input$nprim) # OBSERVABLE[t-1] --> UNOBSERVABLE[t]
input$p=rep(0.3,input$nprim)## CAPTURE PROBABILITY
inputs<-input
ch<-sim_ch(inputs=input)
```

```{r}
library(RMark)
rd<-process.data(data=ch$ch, 
    model="Robust", 
    time.intervals=ch$occs)
S=list(formula=~1)# SURVIVAL
# SHARE = TRUE TO SET C = P
p=list(formula=~1,share=TRUE)# CAPTURE PROBABILITY
f0<- list(formula=~time) # NUMBER NOT ENCOUNTERED
GammaDoublePrime=list(formula=~1,share=TRUE)
GammaPrime=list(formula=~1)
fit<-mark(data = rd, 
    model = "Robust", 
    time.intervals=time.intervals,
    filename="rd",
    model.parameters=list(
        S=S,
        GammaDoublePrime=GammaDoublePrime,
        # GammaPrime=GammaPrime, # not needed when share=TRUE
        p=p),
    threads=2,
    brief=TRUE)
save(fit,file="_output/mark-fit.Rdata")
cleanup(ask=FALSE)
```
```{r,echo=FALSE}
load("_output/mark-fit.Rdata")
```
```{r}
summary(fit)
```

## Estimating demographic rates using Jags

```{r}
T2 <- input$nsec # number of secondary periods per primary period
T <- length(T2) # number of primary periods
ch_array<-array(0,c(nrow(ch$ch_mat),max(T2),T))
secs<-cbind(cumsum(T2),(cumsum(T2)-T2)+1)
for(i in 1:T)
    {
    ch_array[,,i]<- ch$ch_mat[,secs[i,1]:secs[i,2]]
    }


N<- nrow(ch$ch_mat)
n_aug<- N # doubl up population
y<- array(NA,c(N+n_aug, max(T2),T))
y[1:N,,]<-ch_array
y[(N+1):nrow(y),,]<-0# data augmentation
z<- matrix(1,nrow(y),T)
z[1:N,]<-4
dat<-list(
    M = nrow(y),# number of individuals
    T = length(T2),
    T2 = input$nsec,# vector of number of secondary occasions
    y = y) # array of capture histories [ind, sec occasion, primary occasion]
```  


	
```{r}
## initial values
## set for each chain
inits<- function()
	{
	list(lo_omega=rnorm(1,0,1),
        lo_gamma=rnorm(1,0,1),
        lo_phi=rnorm(1,0,0.37),
        lo_p=rnorm(1,0,0.37),z=z)
	}
```

    
    
```{r}
## WHAT PARAMETERS TO KEEP TRACK OF DURING ESTIMATION
params<-c("omega","p","S","lambda","gammaPrime")
```
    
	
```{r}
# THIS WILL ONLY RUN IF YOU HAVE JAGS INSTALLED 
# AND THE R2jags PACKAGE
library(R2jags)
out <- jags(data=dat,
	inits=inits,
	parameters=params,	
	model.file=mod_RD,
	n.chains = 3,	
	n.iter = 250000,	
	n.burnin = 60000, 
	n.thin=2,
	working.directory=getwd())
out  # OUTPUT
sav(out,file="_output/bugs-rd-fit.Rdata")
out$BUGSoutput$mean$N # ESTIMATED NUMBER OF FISH IN POOL
traceplot(out)# LOOK AT TRACEPLOTS FOR CONVERGENCE.
```
